\bf Title here

### 2020-08-12 paper/paper.Rmd
write variable as X_{i2} rather than X_{2i}
elaborate Sec. 2 line search more; move difficulties to Sec. 3; elaborate tour & guided tour more

In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints. The problem consists in optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced spaced given by the problem constraints to either minimize or maximize a function.... Gradient based optimization has been typically used to solve such problems. However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem.

@@

The remainder of the paper is organised as follows. Section \ref{DFO} gives a general overview of optimisation methods, specifically derivative free optimisation. 
Section \ref{tour} introduces the optimisation problem in a specific problem: projection pursuit guided tour and existing derivative free algorithms for solving it. 
Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveil the black-box of optimisation. Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} provides definitions of different diagnostic plots using grammar of graphic and examples. 
Finally, Section \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.

@@

Given an objective function $f$, one way of optimising it is to equate its gradient to zero. In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and Derivative-Free Optimisation (DFO) methods can be useful to approach these problems. One common class of methods in DFO is *Direct-search methods*. Coined by @hooke1961direct, direct search methods don't require any gradient or Hessian information and has gained its popularity through its simplicity in use and reliability in solving complicated practical problems. Depending on whether a random sample is used in the search, this class of methods can be further classified as *stochastic* or *deterministic*.  The stochastic version of the direct-search methods will be the main optimisation procedure analysed in this paper.

[How about adding more details into derivative free methods? ppp]

<!-- The two most common methods in DFO are *direct-search methods* and *model-based methods* and this paper dedicates to the discussion of direct-search methods, which .  -->

<!-- A well-known example of it is the Nelder-Mead algorithm [@nelder1965simplex] and it enjoys the popularity due to its simplicity and reliability. [feel like this sentence can be expanded to include more information. xxx]. -->

- *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large. Also, there are discussions on how to avoid getting trapped in a local optimum in the literature.

- *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible. Stochastic optimisation method could be an alternative to solve these problems.

- *high-dimensional decision variable*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional. Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space.

@@

Coined by @friedman1974projection, projection pursuit detects interesting structures (i.e. clustering, outliers and skewness) in multivariate data via projecting onto lower dimensions. Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis. Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection. A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].

@@

A tour method, guided tour is usually used in conjunction with projection pursuit to interactively explore multivariate data. Tour plays an animation on the projected data which is constructed using the bases outputted by projection pursuit and their interpolations. Guided tour constructs the interpolation using the shortest geodesic path between two outputted bases and we refer readers to @buja2005computational for the mathematical details. Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm and the white frames, which are the interpolations between two blue frames. The tour method has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].

<!-- A tour path includes two major components: a *generator* that generating the target basis that has higher index value than the current basis and an *interpolator* that performing geodesic interpolation between the current and target basis. An illustration modified from [@buja2005computational] vividly depicts the tour path in guided tour. -->

@@

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \text{ and } \mathbf{a}_i \mathbf{a}_j = 0 \}
\end{align}

@@

\begin{equation}
\left[
\begin{array}{c|ccc|cc|cc}
t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
\hline
1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{21}\\
\hline
2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{12}  & V_{22}\\
3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{13}  & V_{23}\\
\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
\hline
\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
\vdots &\vdots &\vdots &\vdots & 2 & 2& \vdots &  \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
\hline
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
\hline
\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{1T}& V_{2T}\\
\hline
\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
\hline
\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{1T^\prime}& V_{2T^\prime}\\
\end{array}
\right]
= 
\left[
\begin{array}{c}
\text{column name} \\
\hline
\text{search (start basis)} \\
\hline
\text{search} \\
\text{search} \\
\vdots \\
\text{search (outputted basis)} \\
\hline
\text{interpolate} \\
\text{interpolate} \\
\vdots \\
\text{interpolate} \\
\hline
\vdots \\
\hline
\text{search} \\
\vdots \\
\text{search (final basis)} \\
\hline
\text{interpolate} \\
\vdots \\
\text{interpolate} \\
\hline
\text{search (no output)} \\
\vdots \\
\text{search (no output)} \\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

>>>>

In an optimization problem the goal is to find the best solution within the space of all feasible solutions which typically is represented by a set of constraints. The problem consists in optimizing an objective function $f: S \rightarrow \Re$ with $S \in \Re^n$ in a reduced space given by the problem constraints to either minimize or maximize a function.... Gradient based optimization has been typically used to solve such problems. However, there are many situations where derivatives of an objective function are unavailable or unreliable and therefore traditional methods based on derivatives are not the best option to solve an optimization problem.

@@

The remainder of the paper is organised as follows. Section \ref{DFO} gives a general overview of optimisation methods, specifically derivative free optimisation. 
Section \ref{tour} introduces the optimisation problem in a specific context: projection pursuit guided tour and existing derivative free algorithms for solving it. 
Section \ref{vis-diag} and \ref{vis-diag-plots} together present the new construction of a visual diagnostic framework that unveils the black-box of optimisation. Section \ref{vis-diag} focuses on motivational ideas and data structure and Section \ref{vis-diag-plots} gives definitions of different diagnostic plots using grammar of graphic and examples. 
Finally, Section \ref{implementation} mentions the R package: ferrn that implements the visual diagnostics described above.

@@

In modern optimisation problems, gradient information can be hard to evaluate or sometimes even impossible and  Derivative-Free Optimisation (DFO) methods can be useful to approach these problems.
A traditional and still reliable in solving complicated practical problems, is *line search method* [@fletcher2013practical]. In the context of a simple one-dimensional problem of finding the $x$ that minimises $f(x)$, line search achieves the goal via an iterative algorithm in the form of Equation \ref{eq:line-search}.

\begin{equation}
x^{(k + 1)} = x^{(k)} + \alpha_k* d^{(k)}
\label{eq:line-search}
\end{equation}

where $d^{(k)}$ is the searching direction in iteration $k$, and $\alpha_k$ is the step-size. Strictly speaking, $\alpha_k$ is chosen by another minimisation of $f(x^{(k)} + \alpha* d^{(k)})$ with respect to $\alpha$ and theoretical results have demonstrated the global convergence of the algorithm when the exact minimisation of $\alpha_k$ is attained [@curry1944method]. In practice, this second minimisation is rarely implemented due to its computational demanding or even the existence of such a minimisation. A more realistic approach is to impose a mandatory decrease in the objective function for each iteration: $f^{(k+1)}> f^{(k)}$ and despite we lose the guarantee on global convergence, this approach turns out to be efficient in practical problems.

@@

Modern development of the line search methods focuses on proposing different computation on the seraching direction: $d^{(k)}$ and various approximations on the step size: $\alpha_k$ catered for practical optimisation problems. The specific problem context we are interested in is called projection pursuit guided tour. 
Projection pursuit and guided tour are two separate methods in exporatory data analysis focusing on different aspects: coined by @friedman1974projection, projection pursuit detects interesting structures (i.e. clustering, outliers and skewness) in multivariate data via low dimensions projection; whilst guided tour is a particular variation in a broader class of data visualisation method called tour.

Let $\mathbf{X}_{n \times p}$ be the data matrix, an n-d projection can be seen as a linear transformation $T: \mathbb{R}^p \mapsto \mathbb{R}^d$ defined by $\mathbf{P} = \mathbf{X} \cdot \mathbf{A}$, where $\mathbf{P}_{n \times d}$ is the projected data and $\mathbf{A}_{p\times d}$ is the projection basis. Define $f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}$ to be an index function that maps the projection basis $\mathbf{A}$ onto an index value $I$, this function is commonly known as the projection pursuit index (PPI) function, or the index function and is used to measure the "interestingness" of a projection. A number of index functions have been proposed in the literature to detect different data structures, including Legendre index [@friedman1974projection], Hermite index [@hall1989polynomial], natural Hermite index [@cook1993projection], chi-square index [@posse1995projection], LDA index [@lee2005projection] and PDA index [@lee2010projection].

@@

Before introducing the guided tour, we shall be familiar with the general tour method [@cook2008grand]. A tour produces animated visualisation of the high dimensional data via rotating low dimension planes. The smoothness of the animation is ensured by computing a series of intermediate planes between two low dimension planes via geodesic interpolation and we refer readers to @buja2005computational for the mathematical details. Iteratively choose different low dimension planes and interpolate between them forms a tour path. Different types of tour methods choose the low dimensional planes differently and we mention two other type of tour that are commonly used. A grand tour selects the planes randomly in the high dimensional space and hence serves as an initial exploration of the data. Manual control allows researches to fine-tuning an existing projection by gradually phase in and out one variable.

Guided tour chooses the planes produced by optimising the projection pursuit index function. Figure \ref{tour-path} shows a sketch of the tour path consisting of the blue frames produced by the projection pursuit optimisation algorithm iteratively and the white frames, which are the interpolations between two blue frames. The tour method has been implemented in the *tourr* package in R, available on the Comprehensive R Archive Network at [https://cran.r-project.org/web/packages/tourr/](https://cran.r-project.org/web/packages/tourr/) [@wickham2011tourrpackage].

@@

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
s.t. &  \langle \mathbf{a}_i, \mathbf{a}_j \rangle = \delta_{ij}, \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}
\end{align}

Where $\delta_{ij}$ is the kronecker delta that takes 1 if $i = j$ and 0 otherwise.

[you wanna give more information about the characteristics of the optimisation problem before introducing existing algorithms ]

---

If the gradient information was available, the decreasing condition is

(feel this section should go after the general introduction of DFO and PPGT)

- *constraint optimisation*: our problem requires the project basis to be orthonormal

- *Multivariate optimisation*: In projection pursuit, the decision variable includes all the entries in the projection matrix, which is high-dimensional. Researchers would be better off if they could understand the relative position of different projection matrix in the high-dimensional space. (multivariate optimisation)

- *optimising non-smooth function*: When the objective function is non-differentiable, derivative information can not be obtained, which means traditional gradient- or Hessian- based methods are not feasible. Stochastic optimisation method could be an alternative to solve these problems. (derivative information is not available)

- *Finding global maximum*: Although finding local maximum is relatively easy with developed algorithms, it is generally hard to guarantee global maximum in a problem where the objective function is complex or the number of decision variables is large. Also, there are discussions on how to avoid getting trapped in a local optimum in the literature. (the exact $\alpha$ that minimised $f(x^{(k)}+ \alpha d^{(k)})$ can be hard to find or sometiems even impossible, imposing certain decrease, how every this doesn't guarantee global convergence)

Computational considerations:

---

@@

\begin{equation}
\left[
\begin{array}{c|ccc|cc|cc}
t & \mathbf{A} & I & S & j &  l  & V_{1} & V_{2}\\
\hline
1 & \mathbf{A}_1 & I_1 & S_1 & 1 & 1 & V_{11} & V_{12}\\
\hline
2 & \mathbf{A}_2 & I_2 & S_2 & 2 & 1  & V_{21}  & V_{22}\\
3 & \mathbf{A}_3 & I_3 & S_3 & 2 & 2  & V_{31}  & V_{32}\\
\vdots & \vdots &\vdots &\vdots  &\vdots & \vdots &\vdots  &\vdots\\
\vdots & \vdots & \vdots &\vdots & 2 & l_2 & \vdots  & \vdots\\
\hline
\vdots &\vdots & \vdots &\vdots & 2  & 1& \vdots & \vdots\\
\vdots &\vdots &\vdots &\vdots & 2 & 2& \vdots &  \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
\vdots &\vdots &\vdots &\vdots & 2 & k_2 &\vdots  & \vdots\\
\hline
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
\hline
\vdots & \vdots & \vdots &\vdots  & J &  1 & \vdots & \vdots \\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
T & \mathbf{A}_T & I_T &S_T  & J &  l_{J} & V_{T1}& V_{T2}\\
\hline
\vdots &\vdots & \vdots &\vdots & J  & 1& \vdots & \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots & \vdots  &\vdots \\
\vdots &\vdots &\vdots &\vdots & J & k_J &\vdots  & \vdots\\
\hline
\vdots& \vdots & \vdots & \vdots & J+1 & 1 & \vdots& \vdots\\
\vdots &\vdots &\vdots &\vdots &\vdots & \vdots &\vdots &\vdots \\
T^\prime & \mathbf{A}_{T^\prime} & I_{T^\prime} &S_{T^\prime}  & J+1 &  l_{J+1} & V_{T^\prime 1}& V_{T^\prime 2}\\
\end{array}
\right]
= 
\left[
\begin{array}{c}
\text{column name} \\
\hline
\text{search (start basis)} \\
\hline
\text{search} \\
\text{search} \\
\vdots \\
\text{search (outputted basis)} \\
\hline
\text{interpolate} \\
\text{interpolate} \\
\vdots \\
\text{interpolate} \\
\hline
\vdots \\
\hline
\text{search} \\
\vdots \\
\text{search (final basis)} \\
\hline
\text{interpolate} \\
\vdots \\
\text{interpolate} \\
\hline
\text{search (no output)} \\
\vdots \\
\text{search (no output)} \\
\end{array}
\right]
\label{eq:data-structure}
\end{equation}

### 2020-08-12 paper/paper.tex
write variable as X_{i2} rather than X_{2i}
elaborate Sec. 2 line search more; move difficulties to Sec. 3; elaborate tour & guided tour more

In an optimization problem the goal is to find the best solution within
the space of all feasible solutions which typically is represented by a
set of constraints. The problem consists in optimizing an objective
function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
spaced given by the problem constraints to either minimize or maximize a
function\ldots{}. Gradient based optimization has been typically used to
solve such problems. However, there are many situations where
derivatives of an objective function are unavailable or unreliable and
therefore traditional methods based on derivatives are not the best
option to solve an optimization problem.

@@

Given an objective function \(f\), one way of optimising it is to equate
its gradient to zero. In modern optimisation problems, gradient
information can be hard to evaluate or sometimes even impossible and
Derivative-Free Optimisation (DFO) methods can be useful to approach
these problems. One common class of methods in DFO is
\emph{Direct-search methods}. Coined by \citet{hooke1961direct}, direct
search methods don't require any gradient or Hessian information and has
gained its popularity through its simplicity in use and reliability in
solving complicated practical problems. Depending on whether a random
sample is used in the search, this class of methods can be further
classified as \emph{stochastic} or \emph{deterministic}. The stochastic
version of the direct-search methods will be the main optimisation
procedure analysed in this paper.

{[}How about adding more details into derivative free methods? ppp{]}

\begin{itemize}
\item
  \emph{Finding global maximum}: Although finding local maximum is
  relatively easy with developed algorithms, it is generally hard to
  guarantee global maximum in a problem where the objective function is
  complex or the number of decision variables is large. Also, there are
  discussions on how to avoid getting trapped in a local optimum in the
  literature.
\item
  \emph{optimising non-smooth function}: When the objective function is
  non-differentiable, derivative information can not be obtained, which
  means traditional gradient- or Hessian- based methods are not
  feasible. Stochastic optimisation method could be an alternative to
  solve these problems.
\item
  \emph{computation speed}: The optimisation procedure needs to be fast
  to compute since the tour produces real-time animation of the
  projected data.
\item
  \emph{consistency result in stochastic optimisation}: In stochastic
  algorithm, researchers usually set a seed to ensure the algorithm
  produces the same result for every run. This practice supports
  reproducibility, while less efforts has been made to guarantee
  different seeds will provide the same result.
\item
  \emph{high-dimensional decision variable}: In projection pursuit, the
  decision variable includes all the entries in the projection matrix,
  which is high-dimensional. Researchers would be better off if they
  could understand the relative position of different projection matrix
  in the high-dimensional space.
\item
  \emph{role of interpolation in PP optimisation}: An optimisation
  procedure usually involves iteratively finding projection bases that
  maximise the index function, while tour requires geodesic
  interpolation between these bases to produce a continuous view for the
  users. It would be interesting to see if the interpolated bases could,
  in reverse, help the optimisation reach faster convergence.
\end{itemize}

@@

Coined by \citet{friedman1974projection}, projection pursuit detects
interesting structures (i.e.~clustering, outliers and skewness) in
multivariate data via projecting onto lower dimensions. Let
\(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection can be
seen as a linear transformation \(T: \mathbb{R}^p \mapsto \mathbb{R}^d\)
defined by \(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{P}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection basis. Define
\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projection basis \(\mathbf{A}\) onto an index
value \(I\), this function is commonly known as the projection pursuit
index (PPI) function, or the index function and is used to measure the
``interestingness'' of a projection. A number of index functions have
been proposed in the literature to detect different data structures,
including Legendre index \citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial}, natural Hermite index
\citep{cook1993projection}, chi-square index
\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
index \citep{lee2010projection}.

@@

A tour method, guided tour is usually used in conjunction with
projection pursuit to interactively explore multivariate data. Tour
plays an animation on the projected data which is constructed using the
bases outputted by projection pursuit and their interpolations. Guided
tour constructs the interpolation using the shortest geodesic path
between two outputted bases and we refer readers to
\citet{buja2005computational} for the mathematical details. Figure
\ref{tour-path} shows a sketch of the tour path consisting of the blue
frames produced by the projection pursuit optimisation algorithm and the
white frames, which are the interpolations between two blue frames. The
tour method has been implemented in the \emph{tourr} package in R,
available on the Comprehensive R Archive Network at
\url{https://cran.r-project.org/web/packages/tourr/}
\citep{wickham2011tourrpackage}.

@@

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
s.t. & \mathcal{A} = \{ \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}: \lVert \mathbf{a}_i \rVert = 1 \text{ and } \mathbf{a}_i \mathbf{a}_j = 0 \}
\end{align}

>>>>

In an optimization problem the goal is to find the best solution within
the space of all feasible solutions which typically is represented by a
set of constraints. The problem consists in optimizing an objective
function \(f: S \rightarrow \Re\) with \(S \in \Re^n\) in a reduced
space given by the problem constraints to either minimize or maximize a
function\ldots{}. Gradient based optimization has been typically used to
solve such problems. However, there are many situations where
derivatives of an objective function are unavailable or unreliable and
therefore traditional methods based on derivatives are not the best
option to solve an optimization problem.

@@

In modern optimisation problems, gradient information can be hard to
evaluate or sometimes even impossible and Derivative-Free Optimisation
(DFO) methods can be useful to approach these problems. A traditional
and still reliable in solving complicated practical problems, is
\emph{line search method} \citep{fletcher2013practical}. In the context
of a simple one-dimensional problem of finding the \(x\) that minimises
\(f(x)\), line search achieves the goal via an iterative algorithm in
the form of Equation \ref{eq:line-search}.

\begin{equation}
x^{(k + 1)} = x^{(k)} + \alpha_k* d^{(k)}
\label{eq:line-search}
\end{equation}

where \(d^{(k)}\) is the searching direction in iteration \(k\), and
\(\alpha_k\) is the step-size. Strictly speaking, \(\alpha_k\) is chosen
by another minimisation of \(f(x^{(k)} + \alpha* d^{(k)})\) with respect
to \(\alpha\) and theoretical results have demonstrated the global
convergence of the algorithm when the exact minimisation of \(\alpha_k\)
is attained \citep{curry1944method}. In practice, this second
minimisation is rarely implemented due to its computational demanding or
even the existence of such a minimisation. A more realistic approach is
to impose a mandatory decrease in the objective function for each
iteration: \(f^{(k+1)}> f^{(k)}\) and despite we lose the guarantee on
global convergence, this approach turns out to be efficient in practical
problems.

@@

Modern development of the line search methods focuses on proposing
different computation on the seraching direction: \(d^{(k)}\) and
various approximations on the step size: \(\alpha_k\) catered for
practical optimisation problems. The specific problem context we are
interested in is called projection pursuit guided tour. Projection
pursuit and guided tour are two separate methods in exporatory data
analysis focusing on different aspects: coined by
\citet{friedman1974projection}, projection pursuit detects interesting
structures (i.e.~clustering, outliers and skewness) in multivariate data
via low dimensions projection; whilst guided tour is a particular
variation in a broader class of data visualisation method called tour.

Let \(\mathbf{X}_{n \times p}\) be the data matrix, an n-d projection
can be seen as a linear transformation
\(T: \mathbb{R}^p \mapsto \mathbb{R}^d\) defined by
\(\mathbf{P} = \mathbf{X} \cdot \mathbf{A}\), where
\(\mathbf{P}_{n \times d}\) is the projected data and
\(\mathbf{A}_{p\times d}\) is the projection basis. Define
\(f: \mathbb{R}^{p \times d} \mapsto \mathbb{R}\) to be an index
function that maps the projection basis \(\mathbf{A}\) onto an index
value \(I\), this function is commonly known as the projection pursuit
index (PPI) function, or the index function and is used to measure the
``interestingness'' of a projection. A number of index functions have
been proposed in the literature to detect different data structures,
including Legendre index \citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial}, natural Hermite index
\citep{cook1993projection}, chi-square index
\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
index \citep{lee2010projection}.

@@

Before introducing the guided tour, we shall be familiar with the
general tour method \citep{cook2008grand}. A tour produces animated
visualisation of the high dimensional data via rotating low dimension
planes. The smoothness of the animation is ensured by computing a series
of intermediate planes between two low dimension planes via geodesic
interpolation and we refer readers to \citet{buja2005computational} for
the mathematical details. Iteratively choose different low dimension
planes and interpolate between them forms a tour path. Different types
of tour methods choose the low dimensional planes differently and we
mention two other type of tour that are commonly used. A grand tour
selects the planes randomly in the high dimensional space and hence
serves as an initial exploration of the data. Manual control allows
researches to fine-tuning an existing projection by gradually phase in
and out one variable.

Guided tour chooses the planes produced by optimising the projection
pursuit index function. Figure \ref{tour-path} shows a sketch of the
tour path consisting of the blue frames produced by the projection
pursuit optimisation algorithm iteratively and the white frames, which
are the interpolations between two blue frames. The tour method has been
implemented in the \emph{tourr} package in R, available on the
Comprehensive R Archive Network at
\url{https://cran.r-project.org/web/packages/tourr/}
\citep{wickham2011tourrpackage}.

@@

\begin{align}
&\arg \max_{\mathbf{A} \in \mathcal{A}} f(\mathbf{X} \cdot \mathbf{A}) \\
s.t. &  \langle \mathbf{a}_i, \mathbf{a}_j \rangle = \delta_{ij}, \forall \mathbf{a}_i, \mathbf{a}_j \in \mathbf{A}
\end{align}

Where \(\delta_{ij}\) is the kronecker delta that takes 1 if \(i = j\)
and 0 otherwise.

{[}you wanna give more information about the characteristics of the
optimisation problem before introducing existing algorithms {]}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

If the gradient information was available, the decreasing condition is

(feel this section should go after the general introduction of DFO and
PPGT)

\begin{itemize}
\item
  \emph{constraint optimisation}: our problem requires the project basis
  to be orthonormal
\item
  \emph{Multivariate optimisation}: In projection pursuit, the decision
  variable includes all the entries in the projection matrix, which is
  high-dimensional. Researchers would be better off if they could
  understand the relative position of different projection matrix in the
  high-dimensional space. (multivariate optimisation)
\item
  \emph{optimising non-smooth function}: When the objective function is
  non-differentiable, derivative information can not be obtained, which
  means traditional gradient- or Hessian- based methods are not
  feasible. Stochastic optimisation method could be an alternative to
  solve these problems. (derivative information is not available)
\item
  \emph{Finding global maximum}: Although finding local maximum is
  relatively easy with developed algorithms, it is generally hard to
  guarantee global maximum in a problem where the objective function is
  complex or the number of decision variables is large. Also, there are
  discussions on how to avoid getting trapped in a local optimum in the
  literature. (the exact \(\alpha\) that minimised
  \(f(x^{(k)}+ \alpha d^{(k)})\) can be hard to find or sometiems even
  impossible, imposing certain decrease, how every this doesn't
  guarantee global convergence)
\end{itemize}

Computational considerations:

\begin{itemize}
\item
  \emph{computation speed}: The optimisation procedure needs to be fast
  to compute since the tour produces real-time animation of the
  projected data.
\item
  \emph{consistency result in stochastic optimisation}: In stochastic
  algorithm, researchers usually set a seed to ensure the algorithm
  produces the same result for every run. This practice supports
  reproducibility, while less efforts has been made to guarantee
  different seeds will provide the same result.
\item
  \emph{role of interpolation in PP optimisation}: An optimisation
  procedure usually involves iteratively finding projection bases that
  maximise the index function, while tour requires geodesic
  interpolation between these bases to produce a continuous view for the
  users. It would be interesting to see if the interpolated bases could,
  in reverse, help the optimisation reach faster convergence.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
